Evolution of ORM Systems: Formal Foundations

### 2013-01-03 doc/description.tex
finished basic operations

\subsection{Software Decomposition}
\begin{gather*}
RemoveClass: Label \times Software \rightarrow Software  \\
\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
\Phi RemoveClass(l, s) = database(s) \bullet dropClass(l) \\
\\ l \in Label, s \in Software 
\\ \\
remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\
\wedge !\: instantiated(c) = \varnothing \wedge !\: associated(c)} {classes(a) \setminus c } \\
\end{gather*}
\begin{gather*}
dropClass(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
\end{gather*}

\begin{gather*}
RemoveProperty: Label \times Label \times Software \rightarrow Software \\
\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
\Phi RemoveProperty(l, l_p, s) =  database(s) \bullet dropProp(l, l_p) \\
RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }\\
l_c, l_p \in Label, a \in Applications 
\end{gather*}
\begin{gather*}
dropProp(p, d) = \begin{cases}
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
} \\ \\
  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
 \end{cases}
\end{gather*}

\begin{gather*}
RemoveAssociation : Label \times Labels \times Software \rightarrow Software \\ 
\Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
\Phi RemoveAssociation(l, l_a, s) = database(s) \bullet dropAssoc(l, l_a) \\
RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
l_c, l_a \in Labels, a \in Applications
\end{gather*}
\begin{gather*}
dropAssoc(l, l_a, d) =  \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = l
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a
 \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
 \\ \\
 \inference{\exists t \in tables(d) : name(t) = l_a
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l
 \\ \wedge \forall r \in data(d) : reference(r) \neq t}{ tables(d) \setminus t}
 \end{cases}
\end{gather*}

\subsection{Software Alternation}
\paragraph{Rename Class}
\begin{gather*}
\rho_{RenC} : Label \times Label \times Application \rightarrow Application \\ 
\rho_{RenC}(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \forall e \in classes(a) : name(e) \neq l_n}{\begin{gathered}
c.name = l_n \wedge \forall f \in associations(e), \\ e \in classes(a), reference(f) = l_c : f.reference = l_n 
\end{gathered}}\\
l_o, l_n \in Label, a \in Applications
\end{gather*}

\paragraph{Rename Property}
\begin{gather*}
\rho_{RenP} : Label \times Label \times Label \times Application \rightarrow Application \\ \\
	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n}{\begin{gathered}
p.name = l_n 
\end{gathered}}\\
l_c, l_o, l_n \in Label, a \in Applications
\end{gather*}

\paragraph{Rename Association}
\begin{gather*}
\rho_{RenA} : Label \times Label \times Label \times Application \rightarrow Application \\
	\rho_{RenP}(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists a \in associations(c) : name(a) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
a.name = l_n 
\end{gathered}}\\
l_c, l_o, l_n \in Label, a \in Applications
\end{gather*}

>>>>

\subsection{Software Decomposition}
\begin{gather*}
RemoveClass: Label \times Software \rightarrow State  \\
\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
\Phi RemoveClass(l, s) = database(s) \bullet dropClass(l) \\
\\ l \in Label, s \in Software 
\\ \\
remClass(l, a) = \inference{\exists c \in classes(a) : name(c) = l \\
\wedge !\: instantiated(c) = \varnothing \wedge !\: associated(c)} {classes(a) \setminus c } \\
\end{gather*}
\begin{gather*}
dropClass(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
\end{gather*}

\begin{gather*}
RemoveProperty: Label \times Label \times State \rightarrow State \\
\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
\Phi RemoveProperty(l, l_p, s) =  database(s) \bullet dropProp(l, l_p) \\
RemoveProperty(l_c, l_p, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_p}{properties(c) \setminus p }\\
l_c, l_p \in Label, a \in Applications 
\end{gather*}
\begin{gather*}
dropProp(p, d) = \begin{cases}
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
} \\ \\
  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
 \end{cases}
\end{gather*}

\begin{gather*}
RemoveAssociation : Label \times Labels \times State \rightarrow State \\ 
\Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
\Phi RemoveAssociation(l, l_a, s) = database(s) \bullet dropAssoc(l, l_a) \\
RemoveAssociation(l_c, l_a, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists f \in associations(c) : name(f) = l_a }{associations(c) \setminus f} \\
l_c, l_a \in Labels, a \in Applications
\end{gather*}
\begin{gather*}
dropAssoc(l, l_a, d) =  \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = l
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a
 \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
 \\ \\
 \inference{\exists t \in tables(d) : name(t) = l_a
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l
 \\ \wedge \forall r \in data(d) : reference(r) \neq t}{ tables(d) \setminus t}
 \end{cases}
\end{gather*}

\subsection{Software Alternation}
\paragraph{Rename Class}
\begin{gather*}
RenameClass : Label \times Label \times State \rightarrow State \\ 
\Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterClass(l_o, l_n) \\ \\
renClass:  Label \times Label \times Application \rightarrow Application \\
renClass(l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_o \\ \wedge \forall e \in classes(a) : name(e) \neq l_n}{\begin{gathered}
c.name = l_n \wedge \forall e \in classes(a) \\ \wedge \forall (f \in associations(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
\end{gathered}}\\ \\
alterTable:  Label \times Label \times Database \rightarrow Database \\
alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
\end{gathered}}\\
l_o, l_n \in Label, s \in States, a \in Applications, d \in Databases
\end{gather*}

\paragraph{Rename Property}
\begin{gather*}
RenameProperty : Label \times Label \times Label \times State \rightarrow State \\ \\
\Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
\Phi(RenameProperty(l_c, l_o, l_n, s)) = database(s) \bullet alterColumn(l_c, l_o, l_n) \\\\
renProp : Label \times Label \times Label \times Application \rightarrow Application \\
	renProp(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists p \in properties(c) : name(p) = l_o \\ \wedge \forall e \in properties(c) : name(e) \neq l_n}{\begin{gathered}
p.name = l_n 
\end{gathered}}\\\\
alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
	alterColumn(l_c, l_o, l_n, d) = \begin{cases}
 \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_o \\ \wedge \forall e \in columns(t) : name(e) \neq l_n}{\begin{gathered}
c.name = l_n 
\end{gathered}}
\\ \\
\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
t.name = l_n 
\end{gathered}}
 \end{cases}
 \\
l_c, l_o, l_n \in Label, a \in Applications, s \in States, d \in Databases
\end{gather*}

\paragraph{Rename Association}
\begin{gather*}
RenameAssociation : Label \times Label \times Label \times State \rightarrow State \\
\Psi(RenameAssocaiation(l_c, l_o, l_n), s) = application(s) \circ renAssoc(l_c, l_o, l_n)\\
\Phi(RenameAssocaiation(l_c, l_o, l_n), s) = database(s) \bullet alterAsspoc(l_c, l_o, l_n)\\\\
renAssoc(l_c, l_o, l_n, a) = \inference{\exists c \in classes(a) : name(c) = l_c \\ \wedge \exists b \in associations(c) : name(b) = l_o \\ \wedge \forall e \in associations(c) : name(e) \neq l_n}{\begin{gathered}
b.name = l_n 
\end{gathered}}\\\\
alterColumn : Label \times Label \times Label \times Database \rightarrow Database \\
	alterAssoc(l_c, l_o, l_n, d) = \begin{cases}
?? \inference{\exists t \in tables(a) : name(t) = l_c \\ \wedge \exists f \in fks(t) : name(c) = l_o \\ \wedge \forall e \in fks(t) : name(e) \neq l_n}{\begin{gathered}
f.name = l_n 
\end{gathered}}
\\ \\
\inference{\exists t \in tables(a) : name(t) = l_o \\ \wedge \forall e \in tables(t) : name(e) \neq l_n}{\begin{gathered}
t.name = l_n 
\end{gathered}}
 \end{cases}
 \\
l_c, l_o, l_n \in Label, a \in Applications, s \in Sates, d \in Databases
\end{gather*}
